Building Sustainable Software - The CSDMS Approach

### 2014-07-14 paper.tex
Cleaned up excess comments.

%\usepackage{multibbl}
%\usepackage{bibtopic}

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
%\usepackage{graphics}
\usepackage[latin9]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{framed,color}
\usepackage{fancybox}
\usepackage{verbatim}

%\setcounter{secnumdepth}{3}
%\setcounter{tocdepth}{3}

@@

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\huttonnote}[1]{ {\textcolor{magenta}    { ***Eric:      #1 }}}
\DeclareRobustCommand{\csdms}{\textsc{csdms}}
\DeclareRobustCommand{\doi}{\textsc{doi}}
\DeclareRobustCommand{\bmi}{\textsc{bmi}}
\DeclareRobustCommand{\wmt}{\textsc{wmt}}

@@

The mission of the Community Surface Dynamics Modeling System (\csdms{})
~\cite{peckham2012component} is to help a diverse modeling community toward
common goals and standards.
%The Community Surface Dynamics Modeling System (\csdms{})
%is an NSF-funded project that began in 2007. Its mission is to help a diverse
%community of surface dynamics model \emph{developers} and model \emph{users}
%work together toward
%common goals and standards.
This effort involves creating:
\begin{itemize}
  \item a repository of source code and metadata for open-source models and
    tools (Section~\ref{sec:organizing})
  \item reusable plug-and-play model components and a framework
        within which these components can be coupled to create new models
        (Section~\ref{sec:reusable})
  \item an efficient and open modeling community through standards (Section
        ~\ref{sec:standards}) and education (Section~\ref{sec:education})
\end{itemize}

%of open-source models. Another part involves converting models to plug-and-play
%components that can be reused via dynamic linking within the \csdms framework to
%create new models.

In building a modeling framework, \csdms{} has leveraged
several existing, well-established and open-source software tools. For example,
\csdms{} uses tools from the Common Component Architecture (CCA)
~\cite{armstrong1999toward} toolchain: \emph{Babel}, and \emph{Bocca}. Babel
provides interoperability between components
written in different languages; it currently supports C, C++, Fortran, Java,
and Python. Bocca helps with creating CCA-compliant components and managing CCA
component projects.
%Ccaffeine is a CCA-compliant framework for linking
%components into new applications. All three tools were designed to enable
%component-based modeling in an HPC environment, but they can also be used on
%desktop systems.

\csdms{} has developed innovative model/component interfaces that promote
model reuse and interoperability,
including the Basic Model Interface (\bmi{} - Section~\ref{sec:bmi}), which
uses
the \csdms{} Standard Names (Section~\ref{sec:standardnames}) and a framework
within which models can be coupled and enhanced with a set of \bmi{}-compatible
service components.
%Component Model Interface (CMI).
The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
graphical interface to this framework and allows users to compose new models
by connecting and configuring components in a simple, browser-based graphical
interface.

%\csdms{} has also developed a web-based client through which users 
%for its modeling framework (the \csdms Web Modeling Tool, WMT, described below)
%that allows users to compose new models by connecting and configuring
%components through a simple, browser-based graphical interface.

@@

%importantly documents the
%submitted code with metadata, smoke-tests codes on the CSDMS High Performance Computing Cluster, maintains a snapshot of submitted source code in a version-control system, and provides easy downloads of the original source code. This repository now contains 172 codes and 56 tools, or 4,074,236 lines of source code (a source code line is defined as any line that is not part of a comment, and is not blank).

%We do know our community values this resource, we track downloads of each individual code and the total number of downloads since November 2008 is >13,218 (as of July 2014).

@@

%\csdms{} is a strong advocate of truly open source software.

%The \doi{} system for numerical models is adopted by
%\csdms as an incentive towards the earth surface processes community, to
%stimulate contributions towards open source software by ensuring recognition
%towards its developers.

@@

%\csdms{} acts as a venue through which open source code of
%earth surface models are promoted, either by hosting or pointing towards open
%source software.

%A Digital Object Identifier (\doi{}) system is implemented to
%guarantee recognition to open source software contributors. \doi{} systems are
%typically used to identify intellectual property in the digital environment.
%It is used principally by publishers, and is an implementation of the Handle
%System for persistent identifiers. The International \doi{} Federation (IDF)
%appoints Registration Agencies who allocate \doi{} prefixes, register \doi{} Names,
%and provide the necessary infrastructure to allow registrants to declare and
%maintain metadata.
%\csdms is among the first software venues to assign \doi{}s to
%open source software. The advantages of adopting a \doi{} system for open source
%software are: 
%\begin{itemize}
%\item Guarantee credit to the software developer
%\item Easily reuse and replicate research as software is directly locatable 
%\item Increase software visibility, \doi{} content is 5 times more likely to
%      deliver active links than content without.
%\item Provide e.g. funding agencies the ability to track usage, so measure
%      impact.
%\end{itemize}

\begin{comment}
\csdms{} requests its \doi{}s through the Integrated Earth Data Applications,
%(\textsc{IEDA}),
a formal Publication Agent of the \doi{} system that operates through the German
National Library of Science and Technology.
\end{comment}

@@

\begin{shaded}
\leftskip 0.25in
\parindent -0.25in
\tt{
Developer, A., Developer, B. (Year of publication). \emph{Name of the model},
Model Version. Identifier.
}
%\emph{ModelDeveloper(s)}, (\emph{PublicationYear}). \emph{ModelName},
%\emph{ModelVersion}. \emph{Identifier}.
\end{shaded}

@@

%The introduction of the
%\csdms Standard Names provides a practical and elegant solution to the semantic
%mediation problem and Babel solves the language interoperability problem.
%\csdms has also assembled a large repository of surface process models that now
%includes over 171 open-source models and 56 tools. The repository contains
%terrestrial, coastal, marine, hydrological, carbonate and atmospheric models.
%Geodynamic, atmospheric and ecosystem models will be added over the next few
%years.

%So far, 55 of the 171 models have been wrapped for use as reusable
%plug-and-play components (although only 6 are currently visible in the new WMT).

@@

%\huttonnote{Taken from the SSI proposal. I need to rewrite this.}

@@

Most models require input variables and produce output variables. For model
components to be reusable and interoperable (either within the \csdms{}
framework or any framework in general)
%In a component-based modeling framework, like \csdms{},
a set of components become a complete model when every component is able to
obtain the input variables it needs from another component. The \csdms{}
Standard Names were developed to provide a practical solution to this
semantic mediation problem ~\cite{peckham2012component, syvitski2014plug}.
%While the CF Convention Standard Names, which were
%introduced in the domain of the ocean and atmosphere model domains, have
%somewhat overlapping goals,
The \csdms{} Standard Names provide a comprehensive set of naming rules and
patterns for creating unique labels for model variables that are not specific
to any particular modeling domain.

%The \csdms Standard Names are currently being extended to solid Earth modeling
%domains, including geodynamics, seismology, electromagnetic induction and
%petrology, in the framework of the "Earth System Bridge" EarthCube project led
%by PI Peckham. They also serve as the base for the metadata standard for
%numerical models (MMF) which is currently in development within \csdms.

%We are acutely aware of multiple other efforts to provide cross-domain
%ontologies in the realm of geosciences. PIs Peckham and Kelbert are involved
%with the initial steps in semantic crosswalk development recently initialized
%within the EarthCube community that aim to achieve synergy between these
%currently divergent metadata ontologies.

@@

%\huttonnote{Taken from the SSI proposal. I need to rework this a bit.}

The \csdms{} Web Modeling Tool\footnote{https://csdms.colorado.edu/wmt}
(\wmt{}) is a web
%is the web-based successor to the desktop Component Modeling Tool. \wmt{} 
application that provides an Ajax client-side (the \wmt{}
client\footnote{https://github.com/csdms/wmt-client}) graphical interface
(figure~\ref{fig:wmt_screenshot}) and a RESTful
server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt}) that allows users to build and run coupled Earth system models on
a high-performance computing cluster from a web browser. \wmt{} was designed
with four objectives:
\begin{itemize}

@@

\item \emph{Portability}. \wmt{} has a native JavaScript interface, so it can
      be accessed on any modern web browser
%, including tablet and mobile
      versions of browsers.

@@

%Although WMT is web-based, the building and configuring of a model can be done
%offline. Reconnection is necessary only when saving a model and submitting it
%for a run.
%\section{Reusable Components}
\begin{comment}
\subsection{Software sustainability through expanded functionality}

@@

%\section{The \csdms Modeling Framework}

@@

%\subsection{CSDMS has an educational mission}
%\subsection{Education and Evaluation}

@@

%IF staff habitually evaluates the effectiveness of this tool and uses the feedback to improve the functionality: modeling courses at the University of Colorado, clinics at the International Summer Institute at the University of Minnesota, and clinics at annual CSDMS Annual Meetings are the dominant forums for this testing.

@@

%Focused hands-on clinics at meetings also allowed more than 150 CSDMS members to experience the ease of use of \wmt{}, and try several examples of how to use the tool.

>>>>

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage[latin9]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{framed,color}
\usepackage{fancybox}
\usepackage{verbatim}

@@

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\huttonnote}[1]{ {\textcolor{magenta}    { ***Eric:      #1 }}}

\DeclareRobustCommand{\csdms}{\textsc{csdms}}
\DeclareRobustCommand{\doi}{\textsc{doi}}
\DeclareRobustCommand{\bmi}{\textsc{bmi}}
\DeclareRobustCommand{\wmt}{\textsc{wmt}}

@@

The mission of the Community Surface Dynamics Modeling System (\csdms{})
~\cite{peckham2012component} is to help a diverse modeling community toward
common goals and standards. This effort involves creating:
\begin{itemize}
  \item a repository of source code and metadata for open-source models and
    tools (Section~\ref{sec:organizing})
  \item reusable plug-and-play model components and a framework
        within which they can be coupled to create new models
        (Section~\ref{sec:reusable})
  \item an \emph{efficient} and \emph{open} modeling community through
        standards (Section ~\ref{sec:standards}) and education
        (Section~\ref{sec:education})
\end{itemize}

In building a modeling framework, \csdms{} has leveraged
several existing, well-established and open-source software tools. For example,
\csdms{} uses tools from the Common Component Architecture (CCA)
~\cite{armstrong1999toward} toolchain: \emph{Babel}, and \emph{Bocca}. Babel
provides interoperability between components
written in different languages; it currently supports C, C++, Fortran, Java,
and Python. Bocca helps with creating CCA-compliant components and managing CCA
component projects.

\csdms{} has developed innovative model/component interfaces that promote
model reuse and interoperability,
including the Basic Model Interface (\bmi{} - Section~\ref{sec:bmi}), which
uses
the \csdms{} Standard Names (Section~\ref{sec:standardnames}) and a framework
within which models can be coupled and enhanced with a set of \bmi{}-compatible
service components.
The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
graphical interface to this framework and allows users to compose new models
by connecting and configuring components in a simple, browser-based graphical
interface.

@@

@@

@@

@@

\begin{shaded}
\leftskip 0.25in
\parindent -0.25in
\tt{
Developer, A., Developer, B. (Year of publication). \emph{Name of the model},
Model Version. Identifier.
}
\end{shaded}

@@

@@

@@

Most models require input variables and produce output variables. For model
components to be reusable and interoperable (either within the \csdms{}
framework or any framework in general)
a set of components become a complete model when every component is able to
obtain the input variables it needs from another component. The \csdms{}
Standard Names were developed to provide a practical solution to this
semantic mediation problem ~\cite{peckham2012component, syvitski2014plug}.
The \csdms{} Standard Names provide a comprehensive set of naming rules and
patterns for creating unique labels for model variables that are not specific
to any particular modeling domain.

@@

The \csdms{} Web Modeling Tool\footnote{https://csdms.colorado.edu/wmt}
(\wmt{}) is a web
application that provides an Ajax client-side (the \wmt{}
client\footnote{https://github.com/csdms/wmt-client}) graphical interface
(figure~\ref{fig:wmt_screenshot}) and a RESTful
server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt}) that allows users to build and run coupled Earth system models on
a high-performance computing cluster from a web browser. \wmt{} was designed
with four objectives:
\begin{itemize}

@@

\item \emph{Portability}. \wmt{} has a native JavaScript interface, so it can
      be accessed on any modern web browser

@@

\begin{comment}
\subsection{Software sustainability through expanded functionality}

@@

@@

@@

@@
